---
title: Meaning and Measurement of External Political Efficacy
subtitle: A Replication of @chamberlain2012
running-head: External Political Efficacy Revisited
author: 
  - name: Isaiah
    affiliations:
      - name: University of Maryland at College Park
        department: Department of Government and Politics 
    email: gespinoz@umd.edu
    # orcid: 
    # corresponding: true
# keywords:
abstract: |
  It is not sufficiently clear what is being measured by external political efficacy measurement items. Not only is the conceptual definition of external efficacy conflated with similar concepts, empirical results are too closely related to satisfactorily distinguish external efficacy as a distinct component on par with internal political efficacy. Drawing from the conventional conceptual definition of external efficacy, @chamberlain2012 hypothesized that external efficacy beliefs would be dynamically responsive to changes in the political environment. Results from the study could not reject the null; aggregated external efficacy was apparently unresponsive across time to chosen predictors. Extending his analysis to include years since the original study partially replicates the original findings. Following the same methodology, updated results show that external efficacy in the aggregate is positively associated with presidential approval ratings with a one year lag. Similar to the previous results, statistical significance disappears upon inclusion of consumer sentiment and trust in government. Such results pose a severe challenge to the construst validity of measurement for external political efficacy. However, results from then and now are contestable on the grounds that the method of analysis does not appropriately test the hypotheses, and the predictors of the analysis are untenable as factors of the political environment. I attempt to conduct a fairer test of the same hypothesis through different methods and factors of the political environment.     

editor_options: 
  chunk_output_type: console
---

```{r}
#| label: setup
#| include: false

# set seed for reproducibility
set.seed(12345)

# load packages (those commented out may or may not be necessary)
library(tidyverse)
library(naniar) # data structures, summaries, and visuals for missing data
library(srvyr) # 'dplyr'-Like Syntax for summary statistics of survey data
library(modelsummary)
library(tinytable)

# load some custom functions
source(here::here('utils', 'funs.R'))

# load data in wide and long formats
d <- readr::read_rds(file = "data/replica data/replica_data_wide.rds")
dlong <- readr::read_rds(file = "data/replica data/replica_data_long.rds")


# set `modelsummary` R package global options
# options(modelsummary_factory_html = 'markdown')
# options(modelsummary_factory_latex = 'kableExtra')

# Replication data set

# the `replica` data set is constrained to the data (presumably) available to
# Chamberlain at the time
replica_long <- dlong |> dplyr::filter(year >= 1952 & year <= 2008)

replica <- d |> dplyr::filter(year >= 1952 & year <= 2008)


```

```{=typst}
#show figure.caption: it => [
    #v(-2em)
    #set text(size:  0.95em, style: "oblique")
    #align(left)[
      #block(inset: 2em)[
        #text(weight: "bold")[
          #it.supplement
          #context it.counter.display(it.numbering)
          ]
          #it.separator
          #it.body
          ]
        ]
      ]

```



## Introduction

Critical scrutiny about the meaning and measure of external political efficacy beliefs has stagnated into comfortable waters. While much effort was expended to construct a reliable and valid measurement method for *internal* political efficacy [@craig1990; @craig1982; @niemi1991a], and later research bolstered its validation [@morrell2003], *external* political efficacy has not received the same critical focus despite its questionable validity. In fact, the two survey items used by the American National Election Studies (ANES) to construct the external political efficacy index have persisted relatively unchanged from the original "sense of political efficacy" measurement scale introduced by @campbell1954 in "The Voter Decides". While the two survey items have remained unchanged, the definition of external political efficacy has deviated substantially from that original concept of political efficacy and even seems divorced from the ordinary meaning of 'efficacy'.

The trouble with the meaning and measure of external political efficacy has not gone unnoticed in prior literature, but there's been little in terms of dedicated scrutiny. One notable effort by @chamberlain2012 stands out as a lonely attempt to validate external efficacy according to it conventional conception. Chamberlain posited that according to the definition of external efficacy, we would expect to see dynamic shifts in aggregate levels of external efficacy in accordance with changes in the political environment. Using aggregate data compiled from multiple sources, he conducted a dynamic time-series analysis on aggregate external political efficacy from 1952 to 2008. Specifically, his study utilized a time-series lagged dependent variable ordinary least squares (OLS) model. The dependent variable was the average index values for external efficacy aggregated across time. The independent variable predictors were aggregated averages of presidential approval ratings, the index of consumer sentiment, trust in government, and a 4-year lag of the dependent variable. The hypothesis was that aggregate external efficacy would appear to be dynamically responsive to the chosen predictors. Null results would suggest the measurement of external efficacy is invalid, and require re-consideration of both concept and measurement. Results could not reject null: aggregate external efficacy was unresponsive to the predictors across time.

Chamberlain's study stands as an uncommon way to test the construct validity of external efficacy that isn't reliant upon methods of common factor analysis or structural equation modeling (SEM). And it is appropriate considering that validating the measure of one construct largely depends on the proposed relations with other distinct constructs. Accordingly, Chamberlain's study was intended to test the construct validity of external political efficacy in reference to how it is understood, defined, and measured in political science and public opinion. However, while sensible, the study misses the mark in a few clear ways. Chamberlain's dynamic time series model with lagged dependent variable is insufficient to properly test the validity and congruence between meaning and measurement of external efficacy. The theoretical justification for the method is lacking, the lags for the model appear arbitrary, but most importantly, the chosen predictors of the model are not tenable as factors representative of the political environment.

Despite these flaws, the general idea that Chamberlain proposed was very sensible: if external political efficacy refers to an individual's belief about the *external* political environment, then we would expect to find that aggregate external efficacy beliefs across time shift upon changes to the political environment. Due to the aforementioned flaws in model and method, the theory and hypothesis still stand to be tested.

The aim of this paper is to attempt a fairer test of external political efficacy following Chamberlain's theoretical propositions. I attempt to replicate results from @chamberlain2012 and then extend them to include observations since the original study. I then address problems with the original study concerning the methods employed to test the theory as well as interpretation of results. In addition, I also examine whether predictors of Chamberlain's original model better relate to satisfaction with government or regime contentment. Then, I conduct a separate analysis with different predictors that, hopefully, serves as a more appropriate test. Using data from the 2016, 2020, and 2024 American National Election Studies (ANES), I examine whether external efficacy differs for panel respondents over time as political conditions differ between them and have changed around them. If external efficacy beliefs of the panel respondents are unresponsive to changes to their respective political conditions, then response data from the survey items cannot be validly interpreted as reflective of the construct according to the conventional definition.

## Background

The initial idea of political efficacy arose instrumentally as a concept for researchers aiming to explain political behavior---an explanation of behavior based on, or in, the psychology of the individual. The seminal book titled "The Voter Decides" by @campbell1954 introduced the concept and initial measure of political efficacy which they believed was related to political activity. These authors referred to this variable as "sense of political efficacy" and described it like this,

> "Sense of political efficacy may be defined as the feeling that individual political action does have, or can have, an impact upon the political process, i.e., that it is worth while to perform one's civic duties. It is the feeling that political and social change is possible, and that the individual citizen can play a part in bringing about this change." [@campbell1954, 187]

The original "sense of political efficacy" scale was measured by having respondents select either "agree" or "disagree" to the following survey items:

```{=typst}
#{
  show table.cell: set text(size: 0.85em)
  show table.cell.where(y: 0): set text(style: "oblique", weight: "bold")
  show table.cell.where(x: 3): smallcaps 
  set table(stroke: 0.6pt)
  
  block(
    inset: 8pt,
    table(
      columns: 3,
      stroke: none,
      table.header(
        table.cell(colspan: 3, align: center)[*Sense of Political Efficacy*]
        ),
      table.hline(),
        ["I don't think public officials care much what people like me think."], [Disagree = 1],
        [NOCARE],
        ["The way people vote is the main thing that decides how things are run in this country."], [Agree = 1], [--],
        ["Voting is the only way that people like me can have any say about how the government runs things."], [Disagree = 1], [VOTING],
        ["People like me don't have any say about what the government does."], [Disagree = 1],
        [NOSAY],
        ["Sometimes politics and government seem so complicated that a person like me can't really understand what's going on."], [Disagree = 1], [COMPLEX]
        )
      )
}


```


At this point, political efficacy was evaluated on the basis of its standing as a Guttman-type scale---as a unidimensional and unidirectional scale with a high coefficient of reproducibility[^1]. Although there were initially five items, the second "The way people vote...." item was dropped from the scale due to its relatively larger error compared to the other items[^2]. Except for this second item, disagreement with the rest of the item statements (NOSAY, NOCARE, VOTING, COMPLEX) was was thought to indicate one's "sense of political efficacy". Positive relationship between one's "sense of political efficacy" and political participation was found and often replicated in subsequent studies [@agger1961; @campbell1960; @dahl1961]. This relationship held after controlling for education and a variety of demographic variables, leading the the authors to conclude that "sense of political efficacy" was, "not merely a function of education or any other likely demographic factor" (@campbell1954, 194).

[^1]: Guttman scales are also referred to as cumulative scales, where hierarchical arrangement of items suggests that agreement/disagreement with one statement implies agreement with with preceding statements. So for a multi-item scale of five statements with dichotomous response options $[0,1]$, a score of $4$ is on a scale of $0-5$. The coefficient of reproducibility reflects how closely a set of items reproduce a "perfect scale". From an individual's scale score alone we would be able to reproduce their pattern of responses to the items perfectly. Of course, we'd occasionally reproduce the wrong response pattern for each person, reflecting the percent of error in the scale. So if a scale has 10 percent error, then we would be able to reproduce their response choices 90 percent of the time. Hence, the coefficient of reproducibility, $C_{R} = 1 - \frac{E}{N \times I}$, where $E$ = sum of the observed response minus expected response, $N$ = number of respondents, and $I$ = number of items [See: @guttman1944; @ford1950; @jobling1961].

[^2]: Ironically, the only survey item asking respondents to express whether the voter decides was dropped from the scale introduced in the book entitled "The Voter Decides".

The first intuitive conception of "sense of political efficacy", however, has been retired; a multidimensional conception of political efficacy now prevails. Today, political efficacy is conceived as an individual's attitude situated along two dimensions of belief. In one dimension, *internal political efficacy* refers to "...beliefs about one's own competence to understand and to participate effectively in politics" [@craig1990, 290; @niemi1991a, 1407]. The other dimension, *external political efficacy*, refers to "...beliefs about the responsiveness of governmental authorities and institutions to citizen demands" [@craig1990, 290; @niemi1991a, 1408]. It is said to be a mistake to refer to the two dimensions as comprising a single, cohesive attitudinal construct [@balch1974]. Thus separate sets of survey items have been devised specifically to measure either attitude as a distinct construct on its own.

In order to assess political efficacy beliefs along the internal dimension, researchers have survey participants indicate the extent to which they agree or disagree with at least four statements, where agreement indicates internal efficacy:

```{=typst}
#{
  show table.cell: set text(size: 0.85em)
  show table.cell.where(y: 0): set text(style: "oblique", weight: "bold")
  show table.cell.where(x: 2): smallcaps 
  set table(stroke: 0.6pt)
  
  block(
    inset: 8pt,
    table(
      columns: 2,
      stroke: none,
      table.header(
        table.cell(colspan: 2, align: center)[*Internal Political Efficacy*]
        ),
      table.hline(),
        ["I consider myself well-qualified to participate in politics"], [SELFQUAL],
        ["I feel I have a pretty good understanding of the important political issues facing our country"], [UNDRSTND],
        ["I feel that I could do as good a job in public office as most other people"], [PUBOFF],
        ["I think that I am better informed about politics and government as most people"],
        [INFORMED]
        )
      )
}


```


A separate set of survey items are consistently utilized to estimate the extent of one's external efficacy beliefs, but in this case there are only two statements:


```{=typst}
#{
  show block: set align(center) 
  show table.cell: set text(size: 0.85em)
  show table.cell.where(y: 0): set text(style: "oblique", weight: "bold")
  show table.cell.where(x: 2): smallcaps
  set table(stroke: 0.6pt)
  
  block(
    inset: 8pt,
    table(
      columns: 2,
      stroke: none,
      table.header(
        table.cell(colspan: 2, align: center)[External Political Efficacy]
        ),
      table.hline(),
        ["People like me don't have any say about what the government does."], [NOSAY],
        ["I don't think public officials care much what people like me think."], [NOCARE]
        )
      )
}


```


For these items, disagreement with either statement is thought to imply positive expression of external efficacy. Historically, survey respondents simply stated their agreement or disagreement with item statements above. Today, response options are usually expanded to include "Strongly agree", "Strongly disagree" and a middling option of "Neither agree nor disagree". 

Today, only two items are used to compose the external political efficacy index and these two items remain virtually unchanged from the original "sense of political efficacy" measurement scale. Whatever conceptual ambiguity remains, the two dimensional conception of political efficacy has been generally accepted in the political science of public opinion and behavior. As such, when researchers refer to political efficacy in general terms, they are most often referring to, and including variable measures for, either the internal or external dimensions of political efficacy.

The eventual empirical deconstruction of a "sense of political efficacy" into multiple dimensions of belief relied on various techniques that seemingly coincided with advances in political methodology and survey research. Validation of the internal and external dimensions of belief has been approached in more than a few ways: comparing relationships of item response data with other variables expected to be related in theory, i.e., external validation [@balch1974]; analysis of measurement reliability and stability of survey items, specifically test-retest reliability [@asher1974]; some form of factor analysis [@craig1982; @finkel1985; @shingles1987; @craig1990; @acock1990; @niemi1991a], to include principal components analysis [@muller1970; @bowler2002], and path analysis techniques [@mcpherson1977]; and finally examination of over time change and correlation of item responses in the aggregate [@converse1972; @chamberlain2012]. Of these, the favored method used to validate the two dimensional construction of political efficacy has been factor analysis[^3].

[^3]: Factor analysis is a multivariate statistical technique that falls under dimension reduction methods, alongside Principal Components analysis and multidimensional scaling, and is also used for the exploration of underlying latent structures within a data set. Broadly speaking, factor analysis encompasses two main approaches: Exploratory Factor Analysis (EFA) and Confirmatory Factor Analysis (CFA). EFA attempts to identify any underlying factor structure of a set of observed variables without making any prior assumptions, i.e., *exploring* the factor structure that might underlie given data. Confirmatory factor analysis, in contrast, is theory driven. In CFA, a researcher hypothesizes the number of factors and patterns of factor loadings expected *a priori* (i.e., proposed in theory) and tests the fit of the factor model. In other words, the CFA approach tests the hypothesis that a set of items are associated with one or more specified factors.

### Conceptual Confusion

The conflation between perceived government responsiveness and external efficacy has made valid interpretation of item response data difficult to stomach. It is not clear whether external efficacy is a belief about perceived government responsiveness or a belief about public capacity to influence government officials given conditions of the political environment. On the one hand, the political context is established as the main focus of the belief from the name of the construct itself; external political efficacy concerns beliefs about their own and others political efficacy (capacity) as conditioned by the *external* political environment. We can think of internal efficacy in much the same way; internal efficacy concerns one's beliefs about their own political efficacy conditioned by that "*internal environment*", so to speak. One's subjective assessment of their own political knowledge, awareness, esteem, and whatever other personal quality pertains to their perceived political competence conditions the extent of what we call their internal political efficacy[^5]. 

[^5]: Indeed it may be less confusing to simply refer to internal political efficacy as perceived personal political competence as many authors already conceive it as such [@almond1963; @converse1972; @abravanel1975]. Especially since immediate links to a generalized "sense of personal competence" and "ego strength" were embedded into the original "sense of political efficacy" idea by psychologists, sociologists, and political scientists [@douvan1956; @douvan1958; @lane1959a; @campbell1960]. And most especially given that internal efficacy does not refer to a belief about the efficacy of political action to achieve political ends.

On the other hand, the perceived responsiveness of government takes central emphasize in the conceptual definition of external political efficacy. Even though the conventional definition refers to external efficacy as a belief about responsiveness *to public input*, it is not defined as an evaluation or appraisal on *the efficacy of public input*. Indeed, external efficacy is primarily an evaluation of perceived government responsiveness and much less an appraisal of the efficacy of public demands, if at all. 

Chamberlain conceptualized external efficacy in such a way that conflates government satisfaction with perceived government responsiveness, and further deviates the meaning of external efficacy away from ordinary meanings of both "external" and "efficacy". Although his overall notion that external political efficacy beliefs should be responsive to shifts in the external political environment was intuitive and sensible, the conceptual definition of external efficacy from which he based his theory and expectations reverted to focus on beliefs about government responsiveness, which in turn led him to base his modeling approach on satisfaction with government.


## Data & Methods {#sec-data-methods}

The American National Election Studies (ANES) have included two items---NOSAY and NOCARE---in almost every survey wave since 1952: every four years starting from 1952, then every two years from 1964 until 2004, and again every four years up to the most recent ANES survey iteration in 2024. Item responses are composed into an index of external political efficacy.

Following @chamberlain2012's original study, the dependent variable consists of aggregated average values of the external political efficacy index derived from the American National Election Studies (ANES) Time Series Cumulative Data File.       

I cannot be entirely sure that the data I have matches or even approximates the presidential approval ratings data he employed in the original analysis. I've documented the process I undertook to gather, process, and compile the data elsewhere.

## Replication of Figure 1 from @chamberlain2012

:::{.place arguments='top, scope: "parent", float: true'}

{{< embed notebooks/figure1_replication.qmd#fig-rep1 >}}

:::

In the original article, the Y-axis of @fig-rep1 was not scaled to range from 0 to 100. Note also that the `NOSAY` external efficacy question was not asked in the 1986 ANES; A line for responses to the `NOSAY` question between 1984 and 1988 is drawn implicitly via linear interpolation.

I extend the time-series chart to include the years of ANES data made available since Chamberlain's original study^[At the time of writing, preliminary data for ANES 2024 was available but not yet included in the Cumulative Data File.] 

:::{.place arguments='top, scope: "parent", float: true'}

{{< embed notebooks/figure1_replication.qmd#fig-rep1adj >}}

:::

## Replication of Figure 2 from @chamberlain2012

@fig-rep2 charts plot lines of the three chosen predictors over time

:::{.place arguments='top, scope: "parent", float: true'}

{{< embed notebooks/figure2_replication.qmd#fig-rep2 >}}

:::

The data available to Chamberlain at the time ranged from 1952 to 2008 for Presidential approval ratings (Gallup); 1960 to 2008 for quarterly ICS data (UMich), and of course, 1958 and 1964 to 2008 for trust in government (ANES). @fig-rep2 constrains the data to approximate the data available at the time of the original publication. 

The following chart @fig-rep2adj is expanded to include all years for which data from each source was available^[Note that the Gallup monthly polling data was obtained through The Roper Center's iPoll database.]. In addition, no adjustment was made to the data and the Y-axis was expanded to preserve the full range of each variable. Values of the ICS are not constrained to range on a particular scale

:::{.place arguments='top, scope: "parent", float: true'}

{{< embed notebooks/figure2_replication.qmd#fig-rep2adj >}}

:::

Note the distinctive difference in ICS values between @fig-rep2 and @fig-rep2adj. In a footnote, Chamberlain states,

> "The Index of Consumer Sentiment was rescaled for figure 2 by subtracting 50 from all its values (to approximate a measure that would fit on a 100-point scale)" [@chamberlain2012, 120, n7]

@fig-rep2compare includes both the non-modified and modified versions of yearly average ICS values over time.

:::{.place arguments='top, scope: "parent", float: true'}

{{< embed notebooks/figure2_replication.qmd#fig-rep2compare >}}

:::

The purpose of @fig-rep2 was not to visually compare average values of each variable. However, such a visualization inevitably presents as such. As stated by Chamberlain, the purpose of the chart was to demonstrate that, "...across time, significant variation exists within these variables." [@chamberlain2012, 120]. 

Understandably, average values are difficult to compare across time due to the fact that they are not all measured on the same scale nor are they obtained by the same survey study. The Index of Consumer sentiment is calculated from a particular formula and then averaged by year; the Trust in Government index is a mean score composite of four question items where higher values indicate higher trust in government; and presidential approval ratings are yearly average approval ratings from monthly Gallup polls^[As an aside, I think the decision to re-scale the ICS by subtracting 50 from the mean values for each year presents a chart that is visibly misleading. This visualization of the data as such easily presents a misleading comparison.].



## Replication of Lagged Dependent Variable Model

Recall @chamberlain2012's hypothesis was that, according to the conceptual definition of the construct, aggregate external efficacy would appear to be dynamically responsive to "changes in the political environment" [-@chamberlain2012, 120]. In order to test this hypothesis, he conducted a dynamic time-series analysis on aggregate external political efficacy from 1952 to 2008 by use of a lagged dependent variable OLS model. The dependent variable was the average index values for external efficacy in the aggregate and across time from 1952 to 2008. His independent variable predictors included presidential approval ratings, an index of consumer sentiment, trust in government, and a 4-year lag of the dependent variable external efficacy. Null results would invalidate, or at least challenge, the measurement of external efficacy and require re-consideration of both concept and measurement. Stated a bit more exhaustively, given the conceptual definition of external political efficacy, null results would invalidate the interpretation that results from the two survey items represent the construct external political efficacy.

Chamberlain reported that results could not reject null: aggregate external efficacy was unresponsive to the predictors across time. I attempted to replicate[^7] the same table of results from the original study [see -@chamberlain2012 123].

[^7]: Note that I did not attempt to replicate 'Model 5' from the original publication. In that model, the dependent variable was regressed on congressional approval and a four-year lag of the dependent variable.

:::{.place arguments='top, scope: "parent", float: true'}

```{r}
#| label: tbl-1
#| tbl-cap: "Replica of Table 1 from Chamberlain (2012): 'Lagged Dependent Variable Model of Macro-Level Variables on External Efficacy'. Note that Model 5 of the original study is not included. Also note that Model 4 has 1 less observation compared to original study, which should not be since the data is exactly the same."
#| tbl-cap-location: bottom


# Essentially, this is an attempt to replicate Table 1 in Chamberlain (2012)
# Except, I exclude Congressional approval as another predictor to examine. I
# didn't gather it from Roper.
# Models 1-4 are models for each individual predictor
# models 5 and 6 are models for sets of predictors (excluding Congressional approval)

models <- list(
  "Model 1" = lm(exteff ~ exteff.T4, data = replica), 
  "Model 2" = lm(exteff ~ exteff.T4 + ics.T1, data = replica), 
  "Model 3" = lm(exteff ~ exteff.T4 + presapp.T1, data = replica), 
  "Model 4" = lm(exteff ~ exteff.T4 + trustgov.T2, data = replica), 
  "Model 5" = lm(exteff ~ exteff.T4 + ics.T1 + presapp.T1, data = replica), 
  "Model 6" = lm(exteff ~ exteff.T4 + ics.T1 + presapp.T1 + trustgov.T2, data = replica))


# vector of preferred coefficient names
coef_names <- c(
  "exteff.T4"   =  "External Efficacy (t-4)",
  "ics.T1"      = "Index of Consumer Sentiments (t-1)",
  "presapp.T1"  = "Presidential Approval (t-1)",
  "trustgov.T2" = "Trust in Government (t-2)",
  "(Intercept)"   = "Constant")

# Replication of results from Chamberlain Table 1.
modelsummary::modelsummary(
  models = models,
  output = "tinytable",
  fmt = 3,
  stars = T,
  coef_map = coef_names,
  gof_map = c("nobs", "adj.r.squared"),
  width = c(.3, .1, .1, .1, .1, .1, .1), 
  align = "lcccccc"
) |> 
  tinytable::style_tt(j = 1, align = "l", bold = T, fontsize = 0.85)

# For model 4, the N I have differs from the N in Chamberlain by 1. This shouldn't be the case since the data should be exactly the same

```

:::

@tbl-1 almost replicates the published results from @chamberlain2012, but with minor differences. The regression coefficients and standard errors are not identical[^8], but very close. However the number of observations for Model 4 differs by one for unknown reasons. The ANES data from the Cumulative Data File (CDF) is identical to the data employed in Chamberlain's study, and as such, Models 1 and 4 from @tbl-1 should not differ from the same models in the original publication. After many arduous attempts, I was unable to discern exactly why the two models differed. 

In the original publication, Chamberlain noted that, "Presidential approval came close to significance as the lone predictor ($p = 0.068$) but was weakened considerably when the index [ICS] and trust were included." [-@chamberlain2012, 122]. The case is similar for replication of Model 3 in @tbl-1; a one-year lag of presidential approval is also significant at the 90\% confidence level ($p = 0.0700$).

[^8]: It is very likely that the data for Presidential approval and Index of Consumer Sentiment I managed to gather for this replication effort differs just enough to explain the slight difference in regression coefficients and standard errors from Table 1 of the original study. Since replication data files were not available to me, I manually sourced and compiled all the data myself from each respective source following the information provided in the original study. 


### Incorporating Years since 2008

:::{.place arguments='top, scope: "parent", float: true'}

```{r}
#| label: tbl-2
#| tbl-cap: "Replication of Table 1 results from @chamberlain2012, including years since original publication."
#| tbl-cap-location: bottom

# Table 1 of Chamberlain extended out to 2024 

# remove ics_modified var
d <- d |> dplyr::select(-ics_modified)

# exclude years prior to 1952 and beyond 2024
d <- d |> 
  dplyr::filter(year >= 1952, year <= 2024)

# NOTE: the lagged DV `exteff.T4` refers to the value of external efficacy at
# t-4, which is four years prior. This is not the same as `exteff.L4` which is
# the 4th previously observed value of external efficacy measured by ANES
# counting back from the most recent (2024).
# For year 2024, `exteff.T4` = 28.5, four years prior
# For year 2024, `exteff.L4` = 38.4, four ANES survey iterations/waves prior but 16 years ago.

models2 <- list(
  "Model 1" = lm(exteff ~ exteff.T4, data = d), 
  "Model 2" = lm(exteff ~ exteff.T4 + ics.T1, data = d), 
  "Model 3" = lm(exteff ~ exteff.T4 + presapp.T1, data = d), 
  "Model 4" = lm(exteff ~ exteff.T4 + trustgov.T2, data = d), 
  "Model 5" = lm(exteff ~ exteff.T4 + ics.T1 + presapp.T1, data = d), 
  "Model 6" = lm(exteff ~ exteff.T4 + ics.T1 + presapp.T1 + trustgov.T2, data = d))


# vector of preferred coefficient names
coef_names2 <- c(
  "exteff.T4"   =  "External Efficacy (t-4)",
  "ics.T1"      = "Index of Consumer Sentiments (t-1)",
  "presapp.T1"  = "Presidential Approval (t-1)",
  "trustgov.T2" = "Trust in Government (t-2)",
  "(Intercept)"   = "Constant")

# Replication of results from Chamberlain Table 1.
modelsummary::modelsummary(
  models = models2, 
  output = "tinytable",
  fmt = 3,
  stars = T,
  coef_map = coef_names2,
  gof_map = c("nobs", "adj.r.squared"),
  width = c(.3, .1, .1, .1, .1, .1, .1), 
  align = "lcccccc"
) |> 
  tinytable::style_tt(j = 1, align = "l", bold = T, fontsize = 0.80)


```

:::

Now Model 3 in @tbl-2 shows presidential approval at $t-1$ statistically significant at the 95\% confidence level ($p = 0.0206$), even when a four-year lag of the dependent variable is included in the model. However including one-year lags of the ICS ($t-1$) in Model 5 renders a one-year lag of presidential approval ($t-1$) significant at 90\% confidence ($p = 0.055$). In Model 6, average presidential approval lagged by one-year is statistically insignificant along with all other predictors of the full model.

However, incorporating data from years since 2008 doesn't change anything for Model 4 in @tbl-2 which remains identical to the same model in @tbl-1 despite the fact that at least four ANES surveys have taken place since 2008. This is due to the fact that average trust-in-government is lagged by two-years. This leaves Model 4 with 20 observations in the aggregate covering only the years from 1966 to 2004.

For some reason, Chamberlain decides to lag the dependent variable by four years but lags the trust-in-government predictor by two years. Starting from 1964 up until 2004, both external efficacy and trust in government were measured every two years, but ANES surveys have been conducted in four-year intervals ever since. Justification for the different lags for external efficacy and trust-in-government is not provided by Chamberlain nor is the reasoning obvious. 

Since external efficacy was measured at an irregular interval, the lag of 4 years is difficult to justify in theory. This 4-year lag ignores external efficacy in periods where there was not a general election even though external efficacy was measured every two years from 1964 up until 2004. The idea is that prior levels of external efficacy influence subsequent levels of external efficacy, but a lag of four years skips years where external efficacy and trust in government was measured by ANES surveys during midterm election cycles. That being said, however, a two-year lag of external efficacy also isn't appropriate. Indeed, the aggregated average of external efficacy in 2024 was $25.1$, but given a two-year lag and that value is null. 

Complicating matters is the tenuous distinction between trust-in-government and external efficacy, conceptually, empirically, and in regards to survey measurement. Given the conceptual overlap between trust-in-government and external political efficacy, it has not been well established that the items measuring one are not also capturing the other.  

### Lagged by Prior Observed Value

Since the general idea is that prior values of aggregated external efficacy account for a portion of subsequent values of the same, then it seems reasonable enough to simply regress external efficacy on the prior observed values of itself instead of its value four years prior. This resolves the issue of measurements taken at irregular intervals over time.

I run the same analysis lagging the dependent variable by it's preceding value, i.e., the average index value of the previous ANES survey iteration. Moreover, I lag the average trust-in-government by its preceding value as well for the models that include the aggregated index.

:::{.place arguments='top, scope: "parent", float: true'}

```{r}
#| label: tbl-3
#| tbl-cap: "Replication of Table 1 results from @chamberlain2012, including years since original publication. External efficacy and trust-in-government are lagged by their preceding values, not by time. Average presidential approval and consumer sentiment maintain one-year lags at T-1."
#| tbl-cap-location: bottom

# Incorporating years up to 2024 into the lagged DV model, where the DV is
# lagged by it's preceding value (not t-1). Also, trust in government is also
# lagged by it's preceding value.

models3 <- list(
  "Model 1" = lm(exteff ~ exteff.L1, data = d), 
  "Model 2" = lm(exteff ~ exteff.L1 + ics.T1, data = d), 
  "Model 3" = lm(exteff ~ exteff.L1 + presapp.T1, data = d), 
  "Model 4" = lm(exteff ~ exteff.L1 + trustgov.T2, data = d), 
  "Model 5" = lm(exteff ~ exteff.L1 + ics.T1 + presapp.T1, data = d), 
  "Model 6" = lm(exteff ~ exteff.L1 + ics.T1 + presapp.T1 + trustgov.L1, data = d))

# running the same LDV models except now external efficacy and trust in
# government are lagged by their previously observed values
# vector of preferred coefficient names
coef_names3 <- c(
  "exteff.L1"   =  "External Efficacy (prior value)",
  "ics.T1"      = "Index of Consumer Sentiments (t-1)",
  "presapp.T1"  = "Presidential Approval (t-1)",
  "trustgov.L1" = "Trust in Government (prior value)",
  "(Intercept)"   = "Constant")

# Replication of results from Chamberlain Table 1.
modelsummary::modelsummary(
  models = models3, 
  output = "tinytable",
  fmt = 3,
  stars = T,
  coef_map = coef_names3,
  gof_map = c("nobs", "adj.r.squared"),
  width = c(.3, .1, .1, .1, .1, .1, .1), 
  align = "lcccccc"
) |> 
  tinytable::style_tt(j = 1, align = "l", bold = T, fontsize = 0.80)

```

:::

Finally, much like Table 1 in Chamberlain's analysis, no predictors save for the lagged dependent variable attain statistical significance. If we choose to understand external political efficacy as beliefs about the responsiveness of government to citizen demands, these results provide us with no supporting information that the survey items validly measure the construct. That is, the public's attitudes about the consumer economy and presidential approval last year, nor the most recent measures of trust in government, seem to bear any weight upon survey items purported to function as measures of beliefs about the responsiveness of government and institutions to citizen demands or preferences. To put in other words, the two items---NOSAY and NOCARE---are not measuring what we think they measure. This was Chamberlain's conclusion implied at the end of his research note. And yet, what we think the items measure is semantically, conceptually, or otherwise sensibly unrelated to a coherent conception of external political efficacy. In fact, it isn't unambiguously clear exactly what it is that political scholars aim to measure in the first place with respect to external political efficacy. The conventional understanding of external political efficacy is misconstrued as government responsiveness and attempts at its measurement are evidently malformed. In fact, research published about decade ago presents a persuasive case for maintaining a clear distinction between external political efficacy and perceived government responsiveness [@esaiasson2015].

## Conclusion: The Bigger Problem

*Efficacy* is a noun. Ordinarily, efficacy refers to the capacity to produce some intended result or outcome. *External political efficacy* is a noun (rather, a noun phrase). The conventional meaning of external political efficacy refers to a belief about the responsiveness of government officials and institutions to public demands---or better said as public influence attempts. Juxtaposing the meanings together like this illustrates that external political efficacy has diverged from having a basic semantic connection with the ordinary meaning of 'efficacy'. Although ordinary meanings do not authoritatively prescribe meaning, external political efficacy is a technical term in political science that is divorced from what it would appear to mean on its face. 

So when we talk about the efficacy of some thing or act, we're discussing the maximum effect of some intervention or stimulus on something else. To say that we appraise the efficacy of something or some act is to say that we are evaluating the capacity of the means toward an explicit end. For instance, researchers of medicine may evaluate the efficacy of some treatment or medical intervention toward the ends of treatment---e.g., vaccine efficacy reveals the capacity of a vaccine to prevent a particular infection or disease. When it comes to external political efficacy as it is defined and understood, we are discussing the extent that one believes the government to be responsive to public demands, preferences, interests, or the like. The intended aim is apparently to induce response from government and the supposed intervention is public influence attempts.

The problem is not simply that we're referring to something by the wrong name. The problem is that the conception of the construct is also detached from the items purported to measure it, and results of those measures do not permit an interpretation coherent with the conventional theory of the concept nor with an improved conception of external political efficacy better aligned with the nominal meaning of the word 'efficacy'.

In broader terms, the problem is that political scholars do not have valid ways to ascertain what we want to know. Ultimately, external efficacy items are uninformative in this regard. More broadly, however, the problem extends to internal political efficacy as well. The measures of internal and external political efficacy do not tell us whether one believes that individual political action is worthwhile, whether political action matters, whether and to what extent active political engagement imposes any influence on the actions of governing actors, nor whether public intervention results in political outcomes. The important part is that, taken together or separately, neither external nor internal efficacy measures tell us the maximum effect of individual nor popular political activity from the perspective of the subject. This means that neither are tenable as evaluations of political efficacy.

For interested researchers, the conventional measures of political efficacy do not nor can not inform us as whether the impact of public intervention, if any, would be recognizable as such from the perspective of the public. This point requires emphasis: scholars are unable to evaluate whether an individual, and by extension the mass public, believes in the efficacy of their own political participation. Consequently, scholars are also unable to evaluate whether the public is sensitive to substantial changes to the political environment; in particular to conditions that pertain to the people's capacity to produce political outcomes, to impose influence, or even to participate as political actors. That is, we're unable to ascertain whether the people's political efficacy beliefs are sensitive to the conditions of their political environment because we have not viable methods to assess those political efficacy beliefs in the first place.

In order to resolve the issue of external efficacy, our conceptual understanding of it must first be addressed and scrutinized. We don't know what we're measuring if we do not have a clear, concise, and distinct conception of it (whatever it is). Especially in relation to the supposed contrast, internal efficacy. Whether that relationship is both coherent and valid is not purely an empirical matter that can be resolved through complex measurement methodology or creative use of the methods of statistical inference. We first must be clear about what it is we mean when we refer to this concept in theory, which, in turn, means that the theory of the concept itself must be valid on its own terms. 

We must understand what we want to know, which is fundamental to scientific inquiry. When we select a measurement instrument, it is presumed that we have clearly identified what we'd like to know because of what results the measurement instrument can produce. When I want to know how much I weigh, I use the appropriate measurement instrument that reliably provides me with valid results corresponding to my weight in an appropriate unit metric for (e.g., pounds lbs, kilos, grams).  

Political efficacy beliefs *should* tell us whether the public is sensitive to the conditions in which they are subject. Political efficacy should tell us whether the people believe themselves capable by the means available, not just to influence governors, but to shape conditions of the political environment as governing actors themselves. If the public is not sensitive to changing political conditions, then they would not recognize when they are subjected to authoritarian regimes and oppressive conditions. Indeed, the people may not recognize themselves as subject at all despite the fact. Democracy, then, is nothing more than a utopia of impotence.

:::{.place arguments='top, scope: "parent", float: true'}

## References {.unnumbered}

:::{#refs}

:::

:::

## Appendix {.appendix}

:::{.place arguments='top, scope: "parent", float: true'}

```{r}
#| label: tbl-presterms
#| tbl-cap: "Presidents"
#| tbl-cap-location: bottom

# just a dataframe (tbl) of presidents with the dates in office
pres_terms <- tibble::tribble(
  ~president,   ~serv_start,         ~serv_end,
  "Roosevelt",    "March 4th, 1933",     "April 12th, 1945",
  "Truman",       "April 12th, 1945",    "January 20th, 1953",
  "Eisenhower",   "January 20th, 1953",  "January 20th, 1961",
  "Kennedy",      "January 20th, 1961",  "November 22nd, 1963",
  "Johnson",      "November 22nd, 1963", "January 20th, 1969",
  "Nixon",        "January 20nd, 1969",  "August 9th, 1974",
  "Ford",         "August 9th, 1974",    "January 20th, 1977",
  "Carter",       "January 20th, 1977",  "January 20th, 1981",
  "Reagan",       "January 20th, 1981",  "January 20th, 1989",
  "Bush",         "January 20th, 1989",  "January 20th, 1993",
  "Clinton",      "January 20th, 1993",  "January 20th, 2001",
  "BushJr",       "January 20th, 2001",  "January 20th, 2009",
  "Obama",        "January 20th, 2009",  "January 20th, 2017",
  "Trump",        "January 20th, 2017",  "January 20th, 2021",
  "Biden",        "January 20th, 2017",  "January 20th, 2025") |>
  dplyr::mutate(pres_num = as.character(seq(32, 46, by = 1)), .before = president) |>
  dplyr::mutate(
    across(c(serv_start, serv_end), ~lubridate::mdy(.)),
    start_year = lubridate::year(serv_start),
    end_year = lubridate::year(serv_end),
    term = lubridate::interval(start = serv_start, end = serv_end, tzone = "UTC")) 

colnames(pres_terms) <- c("Number", "President", "Start", "End")

pres_terms |> 
  dplyr::select(Number, President, Start, End) |> 
  tinytable::tt() |> 
  tinytable::style_tt(
    fontsize = 0.80, # fontsize of text in table in em units (0.75*default)
    align = "llcc"
    )






```

:::









