---
title: Title
author:
  - name: Isaiah
    # orcid: 
    corresponding: true
    email: gespinoz@umd.edu
    # roles:
    #   - Primary 
    affiliations:
      - Department of Government and Politics, University of Maryland at College Park
# keywords:
#   - La Palma
#   - Earthquakes
abstract: |
  I conduct a replication and update of an article authored by Chamberlain in 2012 where he conducted an analysis of aggregate external political efficacy over time in relation to particular predictors.    
date: last-modified
date-format: "YYYY-M-D"
citation:
  container-title: Manuscript of External Efficacy Replication
number-sections: false
---

## Introduction

The literature on political efficacy in political science is stagnate. Chamberlain [-@chamberlain2012] was one of a few who focused on external political efficacy and exposed substantive faults in both its conception and measurement. Rather than propose a new measurement or offer a new conceptual understanding, Chamberlain attempted to discern whether this individual-level construct was sensitive to external factors in the environment as would be expected given its conception and measurement. His research showed that aggregate external efficacy was not responsive to particular changes in the political environment. However, immediately brought into question is whether the predictors selected are reflective or representative of the external political environment in the first place. 



```{r}

```

```{r}
```

```{r}
```


## Data & Methods {#sec-data-methods}

## Conclusion

## References {.unnumbered}

:::{#refs}

:::