---
title: "figure2_replication"
---

```{r}

# set seed
set.seed(12345)

# required packages
# library(tidyverse)
# library(srvyr) # 'dplyr'-Like Syntax for summary statistics of survey data

# load replica data in long formats
dlong <- readr::read_rds(here::here("data", 
                                    "replica data",
                                    "replica_data_long.rds"))

```

```{r}

# Replicate data ----------------------------------------------------------

# the `replica` data set is constrained to the data (presumably) available to
# Chamberlain at the time
replica_long <- dlong |> dplyr::filter(year >= 1952 & year <= 2008)


```

This notebook replicates Figure 2 from @chamberlain2012 (See PDF page 121). Figure 2 charts plot lines of the three chosen predictors over time, and was titled "Changes to the Index of Consumer Sentiment, Presidential Approval, and Trust in Government."


```{r}
#| label: fig-rep2
#| fig-cap: "Replication of Figure 2 in @chamberlain2012 'Changes to the Index of Consumer Sentiment, Presidential Approval, and Trust in Government'. Source data: ANES Time Series Cumulative Data File; The Gallup Organization; Survey of Consumers, University of Michigan."
#| fig-cap-location: bottom
#| fig-asp: 0.7
#| out-width: 60%
#| fig-width: 5.571429



# Note that in a footnote (note 7), Chamberlain states,
# "The Index of Consumer Sentiment was rescaled for figure 2 by subtracting 50 from all its values (to approximate a measure that would fit on a 1OO-point scale)"

# exact replication of Figure 2, modifying ICS values in way described by
# Chamberlain (2012)
fig2.replica <- replica_long |>
  # temporarily remove lagged variables from data frame
  dplyr::select(-dplyr::contains("T")) |>
  # drop rows that contain missing values (NA)
  tidyr::drop_na() |>  
  # exclude average values for external efficacy index and ics (not modified)
  dplyr::filter(!variable %in% c("exteff", "ics")) |> 
  ggplot2::ggplot(ggplot2::aes(x = year, y = value, linetype = variable)) +
  ggplot2::geom_line() +
  ggplot2::scale_linetype_manual(
    values = c(
      "ics_modified" = "solid",
      "presapp" = "dashed",
      "trustgov" = "twodash"
    ),
    labels = c(
      "ics_modified" = "Index of Consumer Sentiment",
      "presapp" = "Presidential Approval",
      "trustgov" = "Trust in Government"
    )
  ) +
  ggplot2::scale_x_continuous(
    limits = c(1950, 2010),
    breaks = seq(1950, 2020, by = 10),
    expan = ggplot2::expansion(0.01, 0.01)
  ) +
  ggplot2::scale_y_continuous(
    limits = c(0, 80),
    breaks = seq(0, 100, by = 20),
    # expansion gets rid of padding
    expand = ggplot2::expansion(0.05, 0.05)) +
  ggplot2::labs(
    title = "Changes to Index of Consumer Sentiment, Presidential Approval, and Trust in Government",
    subtitle = "Replication of Figure 2 from Chamberlain (2012)",
    x = "Year",
    y = "Scale"
  ) +
  ggplot2::theme_bw()+
  ggplot2::theme(
    legend.position = "bottom",
    legend.title = ggplot2::element_blank(),
    panel.background = ggplot2::element_blank(),
    panel.grid = ggplot2::element_blank(),
    plot.title = ggplot2::element_text(hjust = 0.5),
    plot.subtitle = ggplot2::element_text(hjust = 0.5)
  )

fig2.replica

```


The data available to Chamberlain at the time ranged from 1952 to 2008 for Presidential approval ratings (Gallup); 1960 to 2008 for quarterly ICS data (UMich), and of course, 1958 and 1964 to 2008 for trust in government (ANES). @fig-rep2 constrains the data to approximate the data available at the time of the original publication. 

The following chart @fig-rep2adj is expanded to include all years for which data from each source was available^[Note that the Gallup monthly polling data was obtained through The Roper Center's iPoll database.]. In addition, no adjustment was made to the data and the Y-axis was expanded to preserve the full range of each variable. Values of the ICS are not constrained to range on a particular scale, but are calculated using a particular formula. More information on calculation of the ICS can be found in the following [link](https://data.sca.isr.umich.edu/technical-docs.php). 

```{r}
#| label: fig-rep2adj
#| fig-cap-location: bottom
#| fig-cap: "Replication of Figure 2 found in @chamberlain2012 extended to include all years for which data was available. Reflects average Values of the Index of Consumer Sentiment, Presidential Approval, and Trust in Government Over Time. ICS values have not been re-scaled and the Y-Axis is expanded to preserve full range of values. Source data: ANES Time Series Cumulative Data File; The Gallup Organization; Survey of Consumers, University of Michigan."
#| fig-asp: 0.7
#| out-width: 60%
#| fig-width: 5.571429


# Extended Replica of Figure 2 --------------------------------------------

# extended replica of Figure 2. Don't re-scale values; don't cut years.
fig2.adj <- dlong |>
  # temporarily remove lagged variables from data frame
  dplyr::select(-dplyr::contains("T")) |>
  # drop rows that contain missing values (NA)
  tidyr::drop_na() |>
  # exclude average values for external efficacy index and ics (modified)
  dplyr::filter(!variable %in% c("exteff", "ics_modified")) |> 
  ggplot2::ggplot(ggplot2::aes(x = year, y = value, linetype = variable)) +
  ggplot2::geom_line() +
  # ggplot2::geom_point(mapping = aes(colour = variable))+
  ggplot2::scale_linetype_manual(
    values = c(
      "ics" = "solid",
      "presapp" = "dashed",
      "trustgov" = "twodash"
    ),
    labels = c(
      "ics" = "ICS",
      "presapp" = "Presidential Approval",
      "trustgov" = "Trust in Government"
    )
  ) +
  ggplot2::scale_x_continuous(
    limits = c(1938, 2025),
    breaks = seq(1930, 2025, by = 5),
    expan = ggplot2::expansion(0.01, 0.01)
  ) +
  ggplot2::scale_y_continuous(
    limits = c(0, 110),
    breaks = seq(0, 110, by = 10),
    # expansion controls padding at limit y-axis limits
    expand = ggplot2::expansion(0.01, 0.01)) +
  ggplot2::labs(
    # caption = stringr::str_wrap("Source data: ANES Time Series Cumulative Data File; The Gallup Organization; Survey of Consumers, University of Michigan.", width = 95),
    x = "Year",
    y = "Average"
  ) +
  ggplot2::theme_bw()+
  ggplot2::theme(
    legend.position = "bottom",
    legend.title = ggplot2::element_blank(),
    axis.text.x.bottom = ggplot2::element_text(angle = 90),
    panel.background = ggplot2::element_blank(),
    panel.grid = ggplot2::element_blank()
  )

fig2.adj


```

Note the distinctive difference in ICS values between @fig-rep2 and @fig-rep2adj. In a footnote, Chamberlain states,

> "The Index of Consumer Sentiment was rescaled for figure 2 by subtracting 50 from all its values (to approximate a measure that would fit on a 1OO-point scale)" [@chamberlain2012, 120, n7]

@fig-rep2compare includes both the non-modified and modified versions of yearly average ICS values over time.

```{r}
#| label: fig-rep2compare
#| fig-cap-location: bottom
#| fig-cap: "Average Values of the Index of Consumer Sentiment, Presidential Approval, and Trust in Government over time. Includes comparison of modified and original average ICS values over time. Source data: ANES Time Series Cumulative Data File; The Gallup Organization; Survey of Consumers, University of Michigan."
#| fig-asp: 0.7
#| out-width: 60%
#| fig-width: 5.571429


# Compare Exact and Adjusted Replicas of Figure 2 -------------------------

# I present a Figure that includes both the unadjusted and adjusted versions
# of yearly average ICS values over time.

# same replica chart of Figure 2, except now includes non-modified version of
# ICS yearly average values
fig2.compare <- replica_long |>
  # temporarily remove lagged variables from data frame
  dplyr::select(-dplyr::contains("T")) |>
  # drop rows that contain missing values (NA)
  tidyr::drop_na() |> 
  # exclude average values for external efficacy index
  dplyr::filter(!variable %in% c("exteff")) |> 
  ggplot2::ggplot(ggplot2::aes(x = year, y = value, linetype = variable, colour = variable)) +
  ggplot2::geom_line() +
  ggplot2::scale_linetype_manual(
    values = c(
      "ics" = "solid",
      "ics_modified" = "solid",
      "presapp" = "dashed",
      "trustgov" = "twodash"
    ),
    labels = c(
      "ics" = "Index of Consumer Sentiment",
      "ics_modified" = "Index of Consumer Sentiment (Modified)",
      "presapp" = "Presidential Approval", 
      "trustgov" = "Trust in Government"
    )
  ) +
  ggplot2::scale_color_manual(
    values = c("ics" = "black", "ics_modified" = "red",
               "presapp" = "azure4", "trustgov" = "darkgray"),
    labels = c("ics" = "Index of Consumer Sentiment",
               "ics_modified" = "Index of Consumer Sentiment (Modified)",
               "presapp" = "Presidential Approval", 
               "trustgov" = "Trust in Government")
  )+
  ggplot2::scale_x_continuous(
    limits = c(1950, 2010),
    breaks = seq(1950, 2020, by = 10),
    expan = ggplot2::expansion(0.01, 0.01)
  ) +
  ggplot2::scale_y_continuous(
    limits = c(0, 110),
    breaks = seq(0, 100, by = 20),
    # expansion gets rid of padding
    expand = ggplot2::expansion(0.05, 0.05)) +
  ggplot2::labs(
    x = "Year",
    y = "Average"
  ) +
  ggplot2::theme_bw()+
  ggplot2::theme(
    legend.position = "bottom",
    legend.title = ggplot2::element_blank(),
    panel.background = ggplot2::element_blank(),
    panel.grid = ggplot2::element_blank(),
    plot.title = ggplot2::element_text(hjust = 0.5),
    plot.subtitle = ggplot2::element_text(hjust = 0.5)
  )

fig2.compare

```

The purpose of @fig-rep2 was not to visually compare average values of each variable. However, such a visualization inevitably presents as such. As stated by Chamberlain, the purpose of the chart was to demonstrate that, "...across time, significant variation exists within these variables." [@chamberlain2012, 120]. 

Understandably, average values are difficult to compare across time due to the fact that they are not all measured on the same scale nor are they obtained by the same survey study. The Index of Consumer sentiment is calculated from a particular formula and then averaged by year; the Trust in Government index is a mean score composite of four question items where higher values indicate higher trust in government; and presidential approval ratings are yearly average approval ratings from monthly Gallup polls. 

As an aside, I think his decision to re-scale the ICS by subtracting 50 from the mean value for each year presents a chart that is visibly misleading. On visual inspection of @fig-rep2 in the original publication, the average values for aggregate consumer sentiment across time appear far lower than the actual average values derived from the un-adjusted data. Transforming the values of the data keeps the data intact, but in this case, the average ICS values for each year were all subtracted by 50 for the sake of fitting the data to range on a scale between 0 to 100. This may be fine but to me seems questionable; the average values across time are arbitrarily lowered on the y-axis which easily presents a misleading comparison of the values across time.



